A preliminary constraint grammar for Russian

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
figure 1 alignment and discussing asymptoteness of fig 3
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60236 72bbbca6-d526-0410-a7d9-f06f51895060

\begin{figure*}
\centering
\begin{alltt}
{\small
\rus{
"<В>"                                                         "<работу>"
        "в" pr                                                     "работа" n f nn sg acc

"<ноябре>"                                                    "<Международный>"
        "ноябрь" n m nn sg prp                                 "международный" adj m an sg nom
                                                                         "международный" adj m nn sg acc
"<1994>"                                        
        "1994" num                                            "<трибунал>"
                                                                         "трибунал" n m nn sg acc
"<года>"                                                           "трибунал" n m nn sg nom
        "год" n m nn sg gen SELECT:r462                             
;        "год" n m nn pl nom fac SELECT:r462       "<по>"
                                                                          "по" pr 
"<в>"                                                             
        "в" pr                                                    "<Руанде>" 
                                                                          "Руанда" np al f nn sg prp 
"<Танзании>"                                                     "Руанда" np al f nn sg dat
        "Танзания" np al f nn pl acc                           
        "Танзания" np al f nn sg prp                          "<.>" 
;        "Танзания" np al f nn pl nom REMOVE:r424        "." sent 
;        "Танзания" np al f nn sg dat REMOVE:r433 
;        "Танзания" np al f nn sg gen REMOVE:r433

"<начал>"
        "начало" n nt nn pl gen
        "начать" vblex perf tv past m sg
;        "начать" vblex perf iv past m sg REMOVE:r769
}}
\end{alltt}
  \caption{Example output from the morphological analyser and constraint grammar for the sentence \rus{В ноябре 1994 
           года в Танзании начал работу Международный трибунал по Руанде.} ``The work of the International Tribunal 
           for Rwanda started in Tanzania in November 1994.'' The input ambiguity is 1.76 readings per word and 
           the output ambiguity is 1.38 readings per word. Recall is 1.0 and precision is 0.72. Figure~\ref{fig:rules} 
           shows the rules that fired for this example sentence.} 
  \label{fig:output}
\end{figure*}

@@

As can be seen from Figure~\ref{fig:curve}, incorporating the constraint grammar 
improves the performance of the HMM tagger, an improvement of nearly 5\% in accuracy,
similar to that reported by \newcite{hulden12} for the same amount of training data. Our 
constraint grammar also has a much lower precision as a result of the ambiguity remaining
in the output.

The final accuracy is below the state of the art for Russian. For instance, 
\newcite{sharoff08lrec-mocky} report a maximum accuracy of 95.28\% using the TnT tagger.
Note, however, that this model was trained on a much larger corpus -- over five million
tokens -- which is not freely available.

>>>>

\begin{figure*}
\centering
\begin{alltt}
{\small
"<\rus{В}>"                                         "<\rus{работу}>"
    "\rus{в}" pr                                        "\rus{работа}" n f nn sg acc

"<\rus{ноябре}>"                                    "<\rus{Международный}>"
    "\rus{ноябрь}" n m nn sg prp                        "\rus{международный}" adj m an sg nom
                                                 "\rus{международный}" adj m nn sg acc
"<1994>"                                        
    "1994" num                                      "<\rus{трибунал}>"
                                                        "\rus{трибунал}" n m nn sg acc
"<\rus{года}>"                                          "\rus{трибунал}" n m nn sg nom
    "\rus{год}" n m nn sg gen SELECT:r462                             
;   "\rus{год}" n m nn pl nom fac SELECT:r462       "<\rus{по}>"
                                                        "\rus{по}" pr 
"<\rus{в}>"                                                             
    "\rus{в}" pr                                    "<\rus{Руанде}>" 
                                                        "\rus{Руанда}" np al f nn sg prp 
"<\rus{Танзании}>"                                      "\rus{Руанда}" np al f nn sg dat
    "\rus{Танзания}" np al f nn pl acc                           
    "\rus{Танзания}" np al f nn sg prp              "<.>" 
;   "\rus{Танзания}" np al f nn pl nom REMOVE:r424      "." sent 
;   "\rus{Танзания}" np al f nn sg dat REMOVE:r433 
;   "\rus{Танзания}" np al f nn sg gen REMOVE:r433

"<\rus{начал}>"
    "\rus{начало}" n nt nn pl gen
    "\rus{начать}" vblex perf tv past m sg
;   "\rus{начать}" vblex perf iv past m sg REMOVE:r769
}
\end{alltt}
  \caption{Example output from the morphological analyser and constraint grammar for the sentence \rus{В ноябре 1994 
           года в Танзании начал работу Международный трибунал по Руанде.} ``The work of the International Tribunal 
           for Rwanda started in Tanzania in November 1994.'' The input ambiguity is 1.76 readings per word and 
           the output ambiguity is 1.38 readings per word. Recall is 1.0 and precision is 0.72. Figure~\ref{fig:rules} 
           shows the rules that fired for this example sentence.} 
  \label{fig:output}
\end{figure*}

@@

As can be seen from Figure~\ref{fig:curve}, incorporating the constraint grammar 
improves the performance of the HMM tagger, an improvement of nearly 5\% in accuracy,
similar to that reported by \newcite{hulden12} for the same amount of training data. 
In Figure~\ref{fig:curve}, it appears that the HMM alone is much more dependent on 
training corpus size than the voting setup, which improves very little between a training
corpus size of 5,000 and 9,000. This seems to indicate that further improvements depend
more on the CG than on training data.

Our constraint grammar also has a much lower precision as a result of the ambiguity remaining
in the output. Similarly, the final accuracy is below the state of the art for Russian. For instance, 
\newcite{sharoff08lrec-mocky} report a maximum accuracy of 95.28\% using the TnT tagger.
Note, however, that this model was trained on a much larger corpus -- over five million
tokens -- which is not freely available.
